### 2018-01-19 includes/experiment.tex
More rewriting and condensing.

    We generated several problem suites with \fuzzer{} that made one solver perform poorly, but not another. We call those suites: \cHard{} and \zHard{}.

    On \cHard{} \cvc{} did poorly, but \us{} did well. The graphs in Figure~\ref{fig:cvc4-hard} show the performance results for both solvers.

    On \zHard{} \cvc{} did well, but \us{} did poorly. The graphs in Figure~\ref{fig:z3str3-hard} show the performance results for both solvers.

    \todo{Add data on Concats-Extracts}

    On Concats-Extracts ...

>>>>

    We generated several problem suites with \fuzzer{} that made one solver perform poorly, but not another. We call those suites: \cHard{} and \zHard{}. On \cHard{} \cvc{} did poorly, but \us{} did well. The graphs in Figures ~\ref{fig:cvc4-hard} and ~\ref{fig:z3str3-hard} show the performance results.

    \todo{Make real graphs}
    \todo{Add data on Concats-Extracts}

### 2018-01-19 includes/introduction.tex
More rewriting and condensing.

    Many SMT solvers that support the theory of strings have been developed over the past decade, and several have matured to the point of being performant and useful in practice. \theSolvers{}\cite{cvc4}\cite{z3str3} are \numSolvers{} such solvers. Moreover a standard input language for problem instances, \smt{}\cite{smt}, has been developed and many solvers support it. To validate their solvers, many authors have relied on hand-crafted input suites or real-world examples from industrial applications \cite{cvc4-tests} \cite{z3str3-tests} \cite{z3str2-tests}. Indeed several suites of industrial and hand-crafted problems have been compiled \cite{kaluza} \cite{kausler} to both verify solvers and benchmark their performance. However though these suites help solver developers debug and verify their solvers, they do not age well. As solvers grow more complex and expressive, fixed benchmarks fail to exercise new behaviours. As a result, more and more solver functionality goes untested or unverified.

    We introduce a modular fuzzer for string solvers, \fuzzer{}, and describe its utility as an exploratory verification tool. We show its effectiveness by presenting defects we were able to find in existing solvers.

>>>>

    Many SMT solvers that support the theory of strings have been developed over the past decade, and several have matured to the point of being performant and useful in practice. \theSolvers{}\cite{cvc4}\cite{z3str3} are \numSolvers{} such solvers. Moreover a standard input language for problem instances, \smt{}\cite{smt}, has been developed and many solvers support it. To validate their solvers, many authors have relied on hand-crafted input suites or real-world examples from industrial applications \cite{cvc4-tests} \cite{z3str3-tests} \cite{z3str2-tests}. Indeed several suites of industrial and hand-crafted problems have been compiled \cite{kaluza} \cite{kausler} to both verify solvers and benchmark their performance. However though these suites help solver developers debug and verify their solvers, they do not age well. As solvers grow more complex and expressive, fixed benchmarks fail to exercise those new behaviours. As a result, more and more solver functionality goes untested or unverified.

    We introduce a modular fuzzer for string solvers, \fuzzer{}, and describe its utility as an exploratory verification tool. We show its effectiveness by presenting defects we were able to find in leading solvers.

### 2018-01-19 includes/problems.tex
More rewriting and condensing.

    We generated several suites of string and regex problems using \fuzzer{}, and they are available at \problemRepo{}. They are described in more detail below.

    \subsubsection{\fuzzer{}-Generated}

        Most of the problem sets in this suite either have a dedicated generator in \generator{}, or are variations of one. This suite contains the following problem sets: \\

@@

        \begin{tabular}{|c|c|}
            \hline
            \textbf{Name}    & \textbf{Description} \\ \hline
            Lengths          & \\ \hline
            Lengths-Concats  & \\ \hline
            Concats          & \\ \hline
            Concats-Extracts & \\ \hline
            Overlaps         & \\ \hline
            Regex            & \\ \hline
            Many-Regexes     & \\ \hline
            Regex-Deep       & \\ \hline
            Regex-Pair       & \\ \hline
            Regex-Lengths    & \\ \hline
        \end{tabular}

    \subsubsection{\fuzzer{}-Transformed}

        Most of the problem sets in this suite were generated by using one root problem and transforming it with \transformer{} into many similar problems. This suite contains the following problem sets: \\

@@

    While it is likely that some of the pre-generated suites will be immediately useful in analysing a solver, we also found that randomisation was helpful in exploratory debugging, and we encourage this pattern as well. For example the following usage can feed random problems to \cvc{} until it returns an error:

@@

    Some simple UNIX scripting can yield other interesting results. For example the following script finds a regex problem on which \us{} times out after 5 seconds:

>>>>

    We generated several suites of string and regex problems using \fuzzer{}, and they are available at \problemRepo{}. The \texttt{generated} suite contains problems that were generated using \generator{}, described below:

@@

    \begin{tabular}{|c|c|}
        \hline
        \textbf{Name}    & \textbf{Description} \\ \hline
        Lengths          & \\ \hline
        Lengths-Concats  & \\ \hline
        Concats          & \\ \hline
        Concats-Extracts & \\ \hline
        Overlaps         & \\ \hline
        Regex            & \\ \hline
        Many-Regexes     & \\ \hline
        Regex-Deep       & \\ \hline
        Regex-Pair       & \\ \hline
        Regex-Lengths    & \\ \hline
    \end{tabular}
    \hfill \\ \\

    The \texttt{transformerd} suite contains problems generated by using one root problem and transforming it with \transformer{} into many similar problems. They are described below:

@@

    \subsection{Patterns}

        While the aforementioned suites are useful in analysing a solver, we also found that randomisation was helpful in exploratory debugging, and we encourage this pattern as well. For example the following usage can feed random problems to \cvc{} until it returns an error:

@@

        Some \unix{} scripting can yield other interesting results. For example the following script finds a regex problem on which \us{} times out after 5 seconds:

### 2018-01-19 includes/stringfuzz.tex
More rewriting and condensing.

    The \fuzzer{} code is organised to be easily extended: all generators and transformers are independent modules. It is also implemented as a Unix ``filter'', to enable easy integration with other tools (including itself). For example, the outputs of generators can be piped directly into transformers, and transformers can be chained to produce a tuned stream of challenging inputs to a solver.

@@

        \item[\texttt{stringfuzzg}] \hfill \\
            This tool generates \smt{} problems. It supports several generators and options that control its output. The generators it implements are: \\

@@

        \item[\texttt{stringfuzzx}] \hfill \\
            This tool transforms \smt{} problems.. It supports several transformers and options that control its output and input. The transformers it implements are: \\

@@

        \item[\texttt{stringstats}] \hfill \\
            This tool takes as input an \smt{} problem, and outputs its properties. The properties are:

            \begin{itemize}
                \item number of string variables
                \item number of string literals
                \item max/median syntactic depth of expressions
                \item max/median length of literals
            \end{itemize}

>>>>

    The \fuzzer{} code is organised to be easily extended: all generators and transformers are independent modules. Its components are also implemented as \unix{} ``filters'', to enable easy integration with other tools (including themselves). For example, the outputs of generators can be piped directly into transformers, and transformers can be chained to produce a tuned stream of challenging inputs to a solver.

@@

        \item[\texttt{stringfuzzg}] \hfill \\
            This tool generates \smt{} problems. It supports several generators and options that specify its output. The generators it implements are: \\

@@

        \item[\texttt{stringfuzzx}] \hfill \\
            This tool transforms \smt{} problems.. It supports several transformers and options that specify its output and input. The transformers it implements are: \\

@@

        \item[\texttt{stringstats}] \hfill \\
            This tool takes as input an \smt{} problem, and outputs its properties, such as: numbers of string variables and literals, max/median syntactic depth of expressions, and max/median length of literals.
